---
title: "**$Statistical$ $Inference$ $Course$ $Project$ - $Part$ $2$**"
author: "$Giovanny$ $Jimenez$ $Gomez$"
date: "$November$ $25,$ $2020$"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Basic inferential data analysis instructions.

### 1. Load the ToothGrowth data and perform some basic exploratory data analyses.

```{r}
set.seed(2018)
data("ToothGrowth")       # Load Data
str(ToothGrowth)           # Structure of data
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
```

### 2. Provide a basic summary of the data.

```{r, fig.width=5,fig.height=2}
summary(ToothGrowth)           # Structure of data
library(ggplot2)              # In order to provide visualization
mean.supp <- split(ToothGrowth$len, ToothGrowth$supp) # Means by supp
sapply(mean.supp, mean)
ggplot(aes(x = supp, y = len), data = ToothGrowth) + geom_boxplot(aes(fill = supp)) + xlab("Application Method") + ylab("Tooth Length [mm]") + ggtitle("Tooth Length by Application Methods") + scale_fill_manual(name = "Application method", values = c("cornflowerblue","coral3"), labels = c("Orange juice (OJ)", "Vitamin C (VC)"))
mean.dose <- split(ToothGrowth$len, ToothGrowth$dose) # Means by dose
sapply(mean.dose, mean)
ggplot(aes(x = dose, y = len), data = ToothGrowth) + geom_boxplot(aes(fill = dose)) + xlab("Dose [mg]") + ylab("Tooth Length [mm]") + ggtitle("Tooth Length by Dose Levels") + scale_fill_manual(name = "Dose level", values = c("coral3", "forestgreen", "cornflowerblue"), labels = c("0.5mg", "1mg", "2mg"))
```

### 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering).

We are comparing tooth growth by supplement using a $t-test$.
```{r}
t.test(len~supp,data=ToothGrowth)
```
Since the $p-value = 0.06 > 0.05$ and the confidence interval $[-0.17,7.57]$ contains zero, we can say that supplement types seems to have no impact on Tooth growth based on this test.

Now, in order to compare tooth growth by dose, we need to look at the different pairs of dose values.
```{r}
# t-test using dose amounts 0.5 and 1.0 [a 0.5 increase in dosage]
ToothGrowth.sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(1.0,0.5))
t.test(len~dose,data=ToothGrowth.sub)
# t-test using dose amounts 1.0 and 2.0 [a 1 increase in dosage]
ToothGrowth.sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(2.0,1.0))
t.test(len~dose,data=ToothGrowth.sub)
# t-test using dose amounts 0.5 and 2.0 [a 1.5 increase in dosage]
ToothGrowth.sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(2.0,0.5))
t.test(len~dose,data=ToothGrowth.sub)
```

In all three cases:

1. The $p-value$ was approximately zero, and $< 0.05$.
2. The confidence interval does not include zero. 
3. Furthermore, the higher the increase in dosage, the smaller the $p-value$ of the test.

As a result, we can assume that the average tooth length increases with an inceasing dose, and therefore the three null hypothesis can be rejected.

### 4. Conclusions and assumptions.

Given the following assumptions:

1. The sample is representative of the population.
2. The distribution of the sample means follows the Central Limit Theorem.

By observing the t-test analysis above, we can conclude that: 
$(a)$ supplement delivery method has no effect on tooth $growth/length$, but
$(b)$ increased dosages do result in increased tooth length.